EGGs are adhesive!

### 2025-06-15 calco/presentazione/pres.tex
tolta biblio

    \emph{E-graphs (EGGS)} \cite{DetlefsNS05,WillseyNWFTP21} are a graphical formalism  for program optimisation  via an efficient implementation of equality saturation.

@@

     They can be  defined as (acyclic) term graphs \cite{corradini2005term,Plu:TGR-ENTCS} with an additional notion of  equivalence on nodes that is closed under the operators of the signature.

@@

\begin{frame}{Aims of this work} 
    \begin{itemize}
        \item Formalize the definition of EGGs in a categorical setting and study their properties. \pause \medskip 
        \item Develop the theory of EGGs rewriting using the Double Pushout (DPO) paradigm \cite{ehrig2006fundamentals,ehrig1999handbook}.
    \end{itemize}
\end{frame}

@@

\begin{frame}{$\mathcal{M}$-adhesive categories}
	A suitable categorical framework for DPO rewriting is given by the notion of \emph{$\mathcal{M}$-adhesive categories} \cite{azzi2019essence,ehrig2014adhesive, lack2005adhesive}.

		\medskip \pause
		Let $\mathcal{M}$ be a class of monos in a category $\X$, containing all isomorphisms, stable under pullbacks and composition. $\X$ is \emph{$\mathcal{M}$-adhesive} if, given a cube below with the hooked arrows in $\mathcal{M}$, the bottom face a pushout and the back faces pullbacks, we have:

@@

\begin{frame}{Term graphs}
A \emph{term graph} \cite{corradini2005term,corradini1997algebraic, Plu:TGR-ENTCS} is a labelled hypergraphs with injective target function. $\tg$ is be the full subcategory of $\hyp_{\Sigma}$ given by them.

>>>>

    \emph{E-graphs (EGGS)} are a graphical formalism  for program optimisation  via an efficient implementation of equality saturation.

@@

     They can be  defined as (acyclic) term graphs with an additional notion of  equivalence on nodes that is closed under the operators of the signature.

@@

\begin{frame}{Aims of this work} 
    \begin{itemize}
        \item Formalize the definition of EGGs in a categorical setting and study their properties. \pause \medskip 
        \item Develop the theory of EGGs rewriting using the Double Pushout (DPO) paradigm.
    \end{itemize}
\end{frame}

@@

\begin{frame}{$\mathcal{M}$-adhesive categories}
	A categorical framework for DPO rewriting is given by the notion of \emph{$\mathcal{M}$-adhesive categories}.
		\medskip \pause
		Let $\mathcal{M}$ be a class of monos in a category $\X$, containing all isomorphisms, stable under pullbacks and composition. $\X$ is \emph{$\mathcal{M}$-adhesive} if, given a cube below with the hooked arrows in $\mathcal{M}$, the bottom face a pushout and the back faces pullbacks, we have:

@@

\begin{frame}{Term graphs}
A \emph{term graph} is a labelled hypergraphs with injective target function. $\tg$ is be the full subcategory of $\hyp_{\Sigma}$ given by them.
